### 2016-01-16 AssignmentIX/Assignment9-Ditz_Schroeder-WS15-16.tex
bonus aufgabe angefangen aber evtl fallen dir ja noch weitere probleme ein, die wir hier aufschreiben koennen

\begin{document}
%\header{Blatt}{Tutor}{Abgabedatum}{Vorlesung}{Bearbeiter}{Semester}{Anzahl Aufgaben}
\header{8}{Alexander Seitz}{14. December 2015}{Bioinformatics I}{\\Jonas Ditz \\\& Benjamin Schroeder}{WS 15/16}{2}

\section*{Theoretical Assignment - \textit{Suffix tree construction and application}}

\subsection*{a - \textit{Suffix tree construction}}
\subsubsection*{i}
If one uses the naive approach to construct the suffix tree of ``CTAGTAGCAG'', the result would look 
like Figure \ref{fig:naiveImpl}.

\begin{figure}[ht]
 \centering
 \subfigure[step 1]{
  \includegraphics[width=0.15\textwidth]{tast1ai/tree_step01.png}
 }
 \subfigure[step 2]{
  \includegraphics[width=0.25\textwidth]{tast1ai/tree_step02.png}
 }
 \subfigure[step 3]{
  \includegraphics[width=0.35\textwidth]{tast1ai/tree_step03.png}
 }
 \subfigure[step 4]{
  \includegraphics[width=0.45\textwidth]{tast1ai/tree_step04.png}
 }
 \subfigure[step 5]{
  \includegraphics[width=0.45\textwidth]{tast1ai/tree_step05.png}
 }
 \subfigure[step 6]{
  \includegraphics[width=0.45\textwidth]{tast1ai/tree_step06.png}
 }
 \subfigure[step 7]{
  \includegraphics[width=0.45\textwidth]{tast1ai/tree_step07.png}
 }
 \subfigure[step 8]{
  \includegraphics[width=0.45\textwidth]{tast1ai/tree_step08.png}
 }
 \subfigure[step 9]{
  \includegraphics[width=0.45\textwidth]{tast1ai/tree_step09.png}
 }
 \subfigure[step 10]{
  \includegraphics[width=0.7\textwidth]{tast1ai/tree_step10.png}
 }
 \subfigure[step 11]{
  \includegraphics[width=0.7\textwidth]{tast1ai/tree_step11.png}
 }
 \caption{All steps of the naive implementation of suffix tree construction for the string ``CTAGTAGCAG''.}
 \label{fig:naiveImpl}
\end{figure}

\subsubsection*{ii}

\subsection*{b - \textit{Main data table of WOTD}}

\subsection*{c - \textit{Suffix tree application}}

\section*{Theoretical Assignment - \textit{Runtime and space complexity of suffix trees}}

\subsection*{a}
Assume a text $T$ of length $n$ with $n$ times the letter ``a''. If one build a suffix tree for $T$ 
using WOTD, each node will have just one c-group with all remaining suffixes in it. So evaluating the 
root node, one has to compute the longest common prefix of $n$ suffixes, of $n-1$ suffixes for the 
second node and so on. In each c-group the shortest suffix is of length $1$, so for each node $\overline{u}$ 
there are just $|R_a\{\overline{u}\}|$ numbers of comparisons.

Since $T$ is of length $n$ this will lead to an overall runtime of $\sum_{i=1}^{n}i = \frac{1}{2}n(n+1)$, 
which is in $O(n^2)$. $\square$

\subsection*{b}
Suffix links are defined as an auxiliary edge that point from branching node $\overline{bw}$ to the branching 
node $\overline{w}$, if it exists and to the root otherwise. If we now defined a suffix link, such that it 
points into the other direction (i.e. from branching node $\overline{w}$ to branching node $\overline{bw}$) gives 
us an efficient method to find maximal unique matches. If we now find a branching node that indicates 
a unique match (for definition see script) and there is a suffix link that point to another branching node, we know that the current 
branching node is not a maximal unique match (because it is not left maximal). So we just have to 
follow the suffix links until we reach a branching node without an outgoing suffix link and we have 
found a maximal unique match.

\subsection*{c}
A relevant suffix is defined as a suffix that leads to a modification of the tree topology, if it is 
inserted.

>>>>

\begin{document}
%\header{Blatt}{Tutor}{Abgabedatum}{Vorlesung}{Bearbeiter}{Semester}{Anzahl Aufgaben}
\header{9}{Alexander Seitz}{18. January 2016}{Bioinformatics I}{\\Jonas Ditz \\\& Benjamin Schroeder}{WS 15/16}{4}

\section*{Theoretical Assignment - \textit{Overlap and Layout}}

\section*{Theoretical Assignment - \textit{Overlap graph figure}}

\section*{Theoretical Assignment - \textit{Computing all overlaps using suffix tree}}

Finding all overlaps is basically a APSP (All-pairs-suffix-prefix) problem. Solving such a problem 
with a suffix tree approach has a time complexity of $O(n + k^2)$ where $n$ is the total length of 
all strings and $k$ is the count of strings. First all reads will be concatenated with a separation 
character between all reads and a terminal character at the end. Now it is rather simple to check, 
whether the prefix of one read matches with the suffix of another read. This can be done using the 
suffix tree since we know that each prefix starts with an separation character (one could put one 
separation character at the beginning of the string to make this statement true) and every suffix 
ends ether with the separation character or the terminal character. Since one has to go trough the 
whole concatenated string and compare the prefix of each read with the suffix of every other read, 
this approach has a total runtime complexity of $O(n+k^2)$.

\section*{Bonus - \textit{Challenges of your Project}}

One of the biggest challenges during our project was the framework we chose, Vaadin. We were not able 
to get it working of all computers we used. So in the end it was just one of us, who was able to work 
on the front-end layer. That slowed us down a lot. If we would start this project again from the scratch, 
Django would be a better choice. This framework comes with a development server and, therefore, would 
maybe make developing together easier. In the end we wasted a lot of time with Eclipse/Vaadin/Eclipse+Vaadin 
problems. Another challenge was the database schema for POODLE. We, especially Benjamin, talked a lot 
with people from Silke Wiesner's group at the MPI for Developmental Biology and that resulted in 
several schema changes during our project time.
